Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-15 surfaces-write-up/surface-write-up.tex
thm on effective, add related remarks to 1pt

\ssec{One point on Projective Space}
\label{ssec:proj-one-point}

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{write}
\end{rem}

@@

Now, we may apply the relations $L_{i, j}^{\vec{v}, \vec{w}}$
to any monomial in the $F_i^{\vec{v}}$'s involving
indices that differ by at most $1$ to produce the canonical form
\begin{align*}
	(F_i^{\vec{v}_{(1)}})^{g_{\vec{v}_{(1)}}} \cdots
	(F_i^{\vec{v}_{(1)}})^{g_{\vec{v}_{(\ell_1)}}}
	(F_{i + 1}^{\vec{w}_{(1)}})^{g_{\vec{w}_{(1)}}} \cdots
	(F_{i + 1}^{\vec{w}_{(\ell_2)}})^{g_{\vec{w}_{(\ell_2)}}}
\end{align*}

@@

\todo{change thm to be inductive for adding pts, giving result on
effective divisors?}
\todo{remark on not doing two points}

>>>>

\ssec{One Hypersurface on Projective Space}
\label{ssec:proj-one-point}

\begin{rem}
\label{rem:proj-veronese}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{write}
\end{rem}

@@

Now, we may apply the relations $L_{i, j}^{\vec{v}, \vec{w}}$
to any monomial in the $F_i^{\vec{v}}$'s involving
indices that differ by at most $1$ to produce the canonical form
\begin{align*}
	(F_i^{\vec{v}_{(1)}})^{g_{\vec{v}_{(1)}}} \cdots
	(F_i^{\vec{v}_{(\ell_1)}})^{g_{\vec{v}_{(\ell_1)}}}
	(F_{i + 1}^{\vec{w}_{(1)}})^{g_{\vec{w}_{(1)}}} \cdots
	(F_{i + 1}^{\vec{w}_{(\ell_2)}})^{g_{\vec{w}_{(\ell_2)}}}
\end{align*}

@@

\begin{rem}
\label{rem:proj-two-points}
Note that we cannot extend this result to the case when
$D$ is supported at two hypersurfaces with arbitrary rational
coefficients in the same manner that O'Dorney does for the
$\bp^1$ case \cite[Section 4]{dorney:canonical}. As will be shown
in Example ~\ref{eg:hyperplane}, the degrees of generation of the
canonical ring of a general $\bq$-divisor containing two
hyperplanes cannot be bounded so tightly. The two-point
$\bp^1$ result leverages particular facts about the structure of
$\bp^1$.
\end{rem}

\begin{rem}
\label{rem:proj-one-point-ind}
Note that the method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + H$
for some hyperplane $H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-points}.
\end{rem}

\begin{thm}
\label{thm:proj-effective}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i H_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $H_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

Then the canonical ring
\[
	R_D := \bigoplus_{d \geq 0} H^0(\bp^m, \lfloor dD \rfloor)
\]

\noindent
is generated in degrees up to $\max_{1 \leq i \leq n}{q_i}$ with
relations generated in degrees up to $2 \max_{1 \leq i \leq n}{q_i}$
\end{thm}

\begin{proof}
First, we apply Remark ~\ref{rem:proj-veronese} to reduce the
problem to letting each $H_i$ be a hyperplane of $\bp^m$.
Then, the proof follows from applying Remark
~\ref{rem:proj-one-point-ind} inductively on a one-hyperplane case
from Theorem ~\ref{them:proj-one-point}.
\end{proof}
